 ---
title: "SD22.8.6"
author: "Shoumit Dey"
date: "17/10/2023"
output: html_document
---
1. Load packages
2. Set paths
```{r message=FALSE}
#Load packages
library(Seurat)
#define study path
save_path <- "Add your path here"
setwd(save_path)

if (!file.exists(paste0(getwd(), "/R"))){
  dir.create(paste0(getwd(), "/R"))
}

save_path <- paste0(getwd(), "/R/")

#Folder where all samples are stored individually within
#a folder name starting with V* within which a sample h5 file
#(named as *sample*.h5) and its 'spatial folder' should reside 
study_path<-"Folder path for where the RAW files are"


#remove patient 404 Cs and 405 Cs
all_samples <-c("402_3C1","402_3C2","402_3L1","402_3R1","403_1C1","403_1C2","403_1L1","403_1L2","013_8C1", "013_8C2", "013_8L1", "013_8R1")
sample_group <- c("lesion","lesion","healthy","healthy","lesion","lesion","healthy","healthy","lesion","lesion","healthy","healthy")

#create sub-directories
dir.create(paste0(save_path, "markers"))

```
Load all .h5 samples
- Change this subsequently to expression matrix and build stats for generic usage
```{r loadSeurath5}

seuratSpatial <- list()
counter<-1
for (sample in all_samples){
  sample_dir <- Sys.glob(paste0(study_path,"V*/*",sample,"/"))
  sample_path <- Sys.glob(paste0(study_path,"V*/*",sample,"/","*.h5"))
  sample_path_file <- tail(strsplit(sample_path, split = "/")[[1]], n = 1)
  #creating a text indicator for sample origin - eg 'key' LC_402_3C1
  key <- paste0("LC_",sample)
  seuratSpatial[[key]]<-Load10X_Spatial(
                                        sample_dir,
                                        filename = sample_path_file,
                                        assay = "Spatial",
                                        slice = key,
                                        filter.matrix = TRUE,
                                        to.upper = FALSE
                                      )
  #creating a text indicator for sample origin - example LC_402_3C1 is 
  #volunteer 1 from leish challenge cohort
  seuratSpatial[[key]]$orig.ident <- paste0("LC_",sample)
  seuratSpatial[[key]]$group <- sample_group[[counter]]
  counter<-counter+1
}

```
Normalise, scale and find variable features withing each individual sample
```{r SCTRansform, warning=FALSE}

#Normalise, scale and find variable features withing each individual sample
# and regress counts based on nCount_Spatial and nFeature_Spatial
for(sample in all_samples){
  key <- paste0("LC_",sample)
  seuratSpatial[[key]] <- SCTransform(seuratSpatial[[key]], assay = "Spatial", verbose = FALSE, vars.to.regress = c("nCount_Spatial", "nFeature_Spatial"))
}

```
Integrate all samples
```{r integrate_workflow_for_spatial}

features <- SelectIntegrationFeatures(object.list = seuratSpatial, nfeatures = 3000)
seuratSpatial <- PrepSCTIntegration(object.list = seuratSpatial, anchor.features = features)

immune.anchors <- FindIntegrationAnchors(object.list = seuratSpatial, normalization.method = "SCT", anchor.features = features)


integrated_LC_cohort1 <- IntegrateData(anchorset = immune.anchors, normalization.method = "SCT")

rm(seuratSpatial)
```
Cluster integrated samples to find common clusters for underlying skin morphology in health and sand-fly bite-related, or infiltration-related clusters such as 
lesional core, epidermis, dermis, sandfly bite site etc.
```{r clustering}
#run PCA
integrated_LC_cohort1 <- RunPCA(integrated_LC_cohort1, assay = "integrated", verbose = FALSE)
#Visualise variability explained per PC
ElbowPlot(integrated_LC_cohort1)
#set pca dimensions to use and 
#resolution for cluster identification
#Please note various combinations were 
#of PCs and resolution was trialed prior
#selecting based on how well 1:n PCs represented
#the overall variability in the data; and how 
#the underlying morphology was representing by 
#clusterings parameters.
dims=15
res=0.3
#identify neighbours in PC 1:dims dimensions
integrated_LC_cohort1 <- FindNeighbors(integrated_LC_cohort1, reduction = "pca", dims = 1:dims)
#Identify clusters (Louvain)
integrated_LC_cohort1 <- FindClusters(integrated_LC_cohort1, verbose = FALSE, resolution = res)
#Run UMAP using in PC 1:dims dimensions
integrated_LC_cohort1 <- RunUMAP(integrated_LC_cohort1, reduction = "pca", dims = 1:dims)
#Run UMAP using in PC 1:dims dimensions
integrated_LC_cohort1 <- RunTSNE(integrated_LC_cohort1, reduction = "pca", dims = 1:dims)


#write proportion per cluster type across all samples
write.csv(prop.table(table(Idents(integrated_LC_cohort1), integrated_LC_cohort1$orig.ident), margin = 2), paste0(save_path, "/markers/", "_ident_proportions_sample.csv"), row.names =TRUE)
#reordering levels
integrated_LC_cohort1$group <- factor(integrated_LC_cohort1$group, levels = c("healthy","lesion"))
#save rds file
saveRDS(integrated_LC_cohort1, paste0(save_path, "partial_dims_",dims,"_res_",res,"LC_cohort1.Rds"))
```
Sub-Cluster integrated samples but those only identified as overlapping the core lesion in above i.e. cluster 2 as saved in partial_dims_15_res_0.3_LC_cohort1.Rds
```{r sub_clustering_the_core_lesion}
#fetch spots identified as 'lesion core' or cluster 2
cluster2only <- subset(integrated_LC_cohort1, idents = 2)
DefaultAssay(cluster2only)<-"integrated"
#select sub-clustering PCs (1:x) and sub-clustering resolution 
subclust_dims <- 10
subclust_res <- 0.2
#run PCA
cluster2only <- RunPCA(cluster2only, verbose = FALSE)
#Visualise variability explained per PC
ElbowPlot(cluster2only)
#identify neighbours in PC 1:subclust_dims dimensions
cluster2only <- FindNeighbors(cluster2only, reduction = "pca", dims = 1:subclust_dims)
#Identify clusters (Louvain)
cluster2only <- FindClusters(cluster2only, verbose = FALSE, resolution=subclust_res)
#Run UMAP using in PC 1:subclust_dims dimensions
cluster2only <- RunUMAP(cluster2only, reduction = "pca", dims = 1:subclust_dims)
#Run t-SNE using in PC 1:subclust_dims dimensions  
cluster2only <- RunTSNE(cluster2only, reduction = "pca", dims = 1:subclust_dims)
#Prepare for downstream DE analysis
cluster2only <- PrepSCTFindMarkers(cluster2only)
#save Seurat object as Rds file
saveRDS(cluster2only,paste0(save_path, "lesion_core_dims_",subclust_dims,"_res_",subclust_res,".Rds"))
```

```{r cyto_genes}
#fetch genes that were used for re-clustering in next step
genes<-rownames(integrated_LC_cohort1)
ccls_ils_tnfs<-c(grep("^CCL", genes,value = T), grep("^CXC", genes,value = T), grep("^TNF", genes,value = T), grep("^IL", genes,value = T), grep("^IFN", genes,value = T))
```
Sub-Cluster integrated samples but those only identified as overlapping the core lesion in above i.e. cluster 2 as saved in partial_dims_15_res_0.3_LC_cohort1.Rds
```{r re_clustering_by_cytokines_for_lesion_core}
#select sub-clustering PCs (1:x) and sub-clustering resolution 
cyto_dims=10
cyto_res=0.5
#creating a new object for clearer flow - please note this can be easily
#achieved in the already loaded object and the new clustering information
#can be stored in new meta data
lesion_cluster_by_cytokines <- cluster2only
#Change assay to SCT
DefaultAssay(lesion_cluster_by_cytokines) <- "SCT"
#Scale data
lesion_cluster_by_cytokines <- ScaleData(lesion_cluster_by_cytokines, features = ccls_ils_tnfs)
#Run PCA
lesion_cluster_by_cytokines <- RunPCA(lesion_cluster_by_cytokines, verbose = TRUE, features = ccls_ils_tnfs)
#Visualise variability explained per PC
ElbowPlot(lesion_cluster_by_cytokines)
#identify neighbours in PC 1:cyto_dims dimensions
lesion_cluster_by_cytokines <- FindNeighbors(lesion_cluster_by_cytokines, reduction = "pca", dims = 1:cyto_dims)
#Identify clusters (Louvain)
lesion_cluster_by_cytokines <- FindClusters(lesion_cluster_by_cytokines, verbose = TRUE, resolution = cyto_res)
#Run UMAP using in PC 1:subclust_dims dimensions
lesion_cluster_by_cytokines <- RunUMAP(lesion_cluster_by_cytokines, reduction = "pca", dims = 1:cyto_dims)
#Run t-SNE using in PC 1:subclust_dims dimensions
lesion_cluster_by_cytokines <- RunTSNE(lesion_cluster_by_cytokines, reduction = "pca", dims = 1:cyto_dims, check_duplicates=F)

lesion_cluster_by_cytokines$cyto_cluster<-paste0("Cyto_",lesion_cluster_by_cytokines$seurat_clusters)
#ordering names
lesion_cluster_by_cytokines$cyto_cluster <- factor(lesion_cluster_by_cytokines$cyto_cluster,levels=c("Cyto_0","Cyto_1","Cyto_2","Cyto_3","Cyto_4","Cyto_5"))
#setting main indetity per spot as cyto_cluster as defined based on cytokine-based clustering 
Idents(lesion_cluster_by_cytokines)<-"cyto_cluster"

saveRDS(lesion_cluster_by_cytokines, paste0(save_path,"lesion_core_cyto_only_dims_",cyto_dims,"_res_",cyto_res,".Rds"))
```
Read RDS if available
```{r read_RDS_if_available}
res=0.3
dims=15
integrated_LC_cohort1 <- readRDS(paste0(save_path, "partial_dims_",dims,"_res_",res,"_LC_cohort1.Rds"))
subclust_dims <- 10
subclust_res <- 0.2
cluster2only <- readRDS(paste0(save_path, "lesion_core_dims_",subclust_dims,"_res_",subclust_res,".Rds"))
cyto_dims=10
cyto_res=0.5
cluster2CytoOnly <- readRDS(paste0(save_path,"lesion_core_cyto_only_dims_",cyto_dims,"_res_",cyto_res,".Rds"))
```


